Neon: Nuclear Norm to Beat Muon

### 2025-04-26 article/muon.tex
Prototype almost done

\usepackage[pdftex,unicode, 
colorlinks=true,
linkcolor = blue]{hyperref}	% нумерование страниц, ссылки!!!!ИМЕННО В ТАКОМ ПОРЯДКЕ СО СЛЕДУЮЩИМ ПАКЕТОМ
%\usepackage[warn]{mathtext}				% Поддержка русского текста в формулах
\usepackage[T1, T2A]{fontenc}			% Пакет выбора кодировки и шрифтов
\usepackage[utf8]{inputenc} 			% любая желаемая кодировка
\usepackage[english]{babel}		% поддержка русского языка
\usepackage{wrapfig}					% Плавающие картинки
\usepackage{amssymb, amsmath}			% стилевой пакет для формул
\usepackage{algorithm}
\usepackage{algorithmic}

@@

\section{Literature review}
TO-DO.
\section{Quality metrics}
\begin{enumerate}
    \item The derivation is theoretically solid
    \item The numerical procedure used to compute a step is grounded and has estimated time overhead (say, in FLOPS)
    \item The code with Neon trains MLP and CNN (and NanoGPT, but it's a bonus) less than 3 times slower than Adam
    \item Instruction of setting the parameters of the algorithm are presented and justified
    \item The announced article has full structure (Abstract, Introduction, Theory, Experiments, Conclusion, Appendix)
    \item If results are positive, it is written with NeurIPS template.
\end{enumerate}

@@

\section{Prototyping phase report}
\begin{enumerate}
    \item Update rule is derived, see idea
    \item Update rule methods are tested: power iteration vs Lanzos (see playground)
    \item NanoGPT is tested on Muon and Adam (add fig). The results are strange, because the methods do not converge at the expected rate (Muon is slower)
    \item Neon, Muon, AdamW and SGD are compared on MLP and CNN (add fig). All methods work correctly, but again there is the problem with which one is the fastest (for now, it's SGD).
\end{enumerate}

\bibliographystyle{unsrt}
\bibliography{muon}

>>>>

\usepackage[pdftex,unicode, 
colorlinks=true,
linkcolor = blue]{hyperref}	% нумерование страниц, ссылки!!!!ИМЕННО В ТАКОМ ПОРЯДКЕ СО СЛЕДУЮЩИМ ПАКЕТОМ
\usepackage{natbib}  % Add natbib for better citation handling
%\usepackage[warn]{mathtext}				% Поддержка русского текста в формулах
\usepackage[T1, T2A]{fontenc}			% Пакет выбора кодировки и шрифтов
\usepackage[utf8]{inputenc} 			% любая желаемая кодировка
\usepackage[english]{babel}		% поддержка русского языка
\usepackage{wrapfig}					% Плавающие картинки
\usepackage{amssymb, amsmath}			% стилевой пакет для формул
\usepackage{algorithm}
\usepackage{algorithmic}

@@

\section{Literature review}
TO-DO

\section{Quality metrics}
\begin{enumerate}
    \item The derivation is theoretically solid
    \item The numerical procedure used to compute a step is grounded and has estimated time overhead (say, in FLOPS)
    \item The code with Neon trains MLP and CNN (and NanoGPT, but it's a bonus) less than 3 times slower than Adam
    \item Instruction of setting the parameters of the algorithm are presented and justified
    \item The announced article has full structure (Abstract, Introduction, Theory, Experiments, Conclusion, Appendix)
    \item If results are positive, it is written with NeurIPS template.
\end{enumerate}

@@

\section{Prototyping phase report}
\begin{enumerate}
    \item Update rule is derived, see idea
    \item Update rule methods are tested: power iteration vs Lanzos (see playground)
    \item Recorded the distribution of singular values of gradients during NanoGPT training (add fig).
    \item NanoGPT is tested on Muon and Adam (add fig). The results are strange, because the methods do not converge at the expected rate (Muon is slower). Besides, Neon doe not converge (add fig).
    \item Neon, Muon, AdamW and SGD are compared on MLP and CNN (see Figures~\ref{fig:mlp_epochs}, \ref{fig:cnn_epochs}, \ref{fig:mlp_time}, and~\ref{fig:cnn_time}). All methods work correctly, but again there is the problem with which one is the fastest (for now, it's SGD).
    \begin{figure}[h!]
        \center{\includegraphics[width=0.8\linewidth]{../figs/mlp24/accuracy_vs_epochs_simple.png}}
        \caption{MLP: self.linear1 = nn.Linear(32*32*3, 512), self.linear2 = nn.Linear(512, 10), self.activ = nn.GELU()}
        \label{fig:mlp_epochs}
    \end{figure}
    \begin{figure}[h!]
        \center{\includegraphics[width=0.8\linewidth]{../figs/mlp24/accuracy_vs_epochs_moderate.png}}
        \caption{CNN: 2 convolutional blocks, 2 fully connected layers, activation + dropout}
        \label{fig:cnn_epochs}
    \end{figure}
    \begin{figure}[h!]
        \center{\includegraphics[width=0.8\linewidth]{../figs/mlp24/accuracy_vs_time_simple.png}}
        \caption{MLP: wallclock time measurements}
        \label{fig:mlp_time}
    \end{figure}
    \begin{figure}[h!]
        \center{\includegraphics[width=0.8\linewidth]{../figs/mlp24/accuracy_vs_time_moderate.png}}
        \caption{CNN: wallclock time measurements}
        \label{fig:cnn_time}
    \end{figure}
\end{enumerate}

\bibliographystyle{plainnat}  % Change to plainnat style
\bibliography{muon}  % Make sure to include the bibliography
